Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based cohort study

### 2024-11-25 manuscript2.Rmd
25/11

Table 1 presents sample characteristics. Most patients were male 1,096 (76%) vs 353 (24%). Of the 354 (42%) patients who were intubated, 241 (17%) patients received mechanical ventilation for 1-7 days, while 113 (7,8%) patients received mechanical ventilation for more than 7 days. The number of patients arriving outside working hours was 1,042 (72%). A total of 1,064 (73%) patients were equal to or less than 7 days in the ICU vs 385 (27%) patients whose ICU-stay exceeded 7 days. A total of 1,080 (75%) patients had an ASA score of 1-2 before arrival to the hospital, and 369 (25%) patients had an ASA score of 3-6 before arrival to the hospital. Amongst our patients, 30-day mortality was at 286 (20%). Of these, 92 patients died within 24 hours, and a total of 144 patients died within 48 hours of hospital arrival. When comparing patients with opportunities for improvement to those without, patients with opportunities for improvement were less likely to require intubation, had higher Revised Trauma Score, experienced shorter time to initial computed tomography scans, spent more days in the ICU, and demonstrated a greater likelihood of surviving 30 days post-injury.

@@

Table 2 illustrates opportunities for improvement divided into categories based on the definitions of opportunity for improvement types during mortality and morbidity reviews at the Karolinska University Hospital in Solna. The most common category of opportunity for improvement in the patient cohort of this study was clinical judgement error, followed by delay in treatment.

@@

The positive association between the Revised Trauma Score and opportunities for improvement, means that among the very severely injured patients admitted to the ICU, patients in a less severe condition on admission were at higher odds of opportunities for improvement. One hypothetical explanation is that patients with higher Revised Trauma Score, indicating relative stability upon arrival, may experience delays in interventions overall. This aligns with findings from previous research conducted on all trauma patients from the Karolinska University Hospital Solna, where it was noted that patients with more stable vital parameters were more likely to have opportunities for improvement [@12].

>>>>

Table 1 presents sample characteristics. Most patients were male 1,096 (76%) vs 353 (24%). Of the 354 (42%) patients who were intubated, 241 (17%) patients received mechanical ventilation for 1-7 days, while 113 (7,8%) patients received mechanical ventilation for more than 7 days. The number of patients arriving outside working hours was 1,042 (72%). A total of 1,064 (73%) patients were equal to or less than 7 days in the ICU vs 385 (27%) patients whose ICU-stay exceeded 7 days. A total of 1,080 (75%) patients had an ASA score of 1-2 before arrival to the hospital, and 369 (25%) patients had an ASA score of 3-6 before arrival to the hospital. Amongst our patients, 30-day mortality was at 286 (20%). Of these, 105 patients died within 24 hours, and a total of 148 patients died within 48 hours of hospital arrival. When comparing patients with opportunities for improvement to those without, patients with opportunities for improvement were less likely to require intubation, had higher Revised Trauma Score, experienced shorter time to initial computed tomography scans, spent more days in the ICU, and demonstrated a greater likelihood of surviving 30 days post-injury.

@@

Table 2 illustrates opportunities for improvement divided into categories based on the definitions of opportunity for improvement types during mortality and morbidity reviews at the Karolinska University Hospital in Solna. The most frequently identified opportunity for improvement in the patient cohort of this study was clinical judgement errors, accounting for 42 cases (29%), followed by delays in treatment, observed in 33 cases (23%).

@@

The positive association between the Revised Trauma Score and opportunities for improvement, means that among the very severely injured patients admitted to the ICU, patients who are more physiologically stable upon admission were at higher odds of opportunities for improvement. One hypothetical explanation is that patients with higher Revised Trauma Score, indicating relative stability upon arrival, may experience delays in interventions overall. Previous research on all trauma patients at the Karolinska University Hospital Solna has demonstrated differences in opportunities for improvement between all trauma patients and those treated in the ICU [@12].
